EldersAIR: manuscript tables/figures

### 2022-07-21 manuscript_tables_figures.Rmd
update to health analysis and tables

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))  %>% 
  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019")))
  left_join(treatments_blind, by = c("area", "home_id_num"))

@@

# Summary stats for character/factor data
```{r}
# summary data for manuscript tables
char_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>%
  #filter(area == "NPT") %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  mutate(total_n = n()) %>% 
  group_by(area) %>% 
  mutate(area_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = n()/total_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(new_var)
demographics_summary
}

char_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>% 
  #filter(area == "NPT") %>% 
  arrange(home, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(home, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  mutate(total_n = n()) %>% 
  group_by(treatment) %>% 
  mutate(treatment_n = n()) %>% 
  group_by(treatment, new_var) %>% 
  mutate(n = n(),
         percent = n()/treatment_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(treatment, new_var)
demographics_summary
}

# distinct by participant
char_funct_total("gender")
char_funct_treatment("gender")
char_funct_total("hispanic")
char_funct_treatment("hispanic")
char_funct_total("race")
char_funct_treatment("race")
char_funct_total("education")
char_funct_treatment("education")
char_funct_total("cohort")
char_funct_treatment("cohort")

@@

# distinct by participant
num_funct_total("age")
num_funct_treatment("age")

@@

# distinct by participant
num_funct_total("bp_sys")
num_funct_treatment("bp_sys")
num_funct_total("bp_dia")
num_funct_treatment("bp_dia")
num_funct_total("bmi")
num_funct_treatment("bmi")
num_funct_total("fvc")
num_funct_treatment("fvc")
num_funct_total("fev1")
num_funct_treatment("fev1")
num_funct_total("fev1_fvc_ratio")
num_funct_treatment("fev1_fvc_ratio")
num_funct_total("fvc_pp")
num_funct_treatment("fvc_pp")
num_funct_total("fev1_pp")
num_funct_treatment("fev1_pp")
num_funct_total("mp_mean_winter")
num_funct_treatment("mp_mean_winter")

@@

>>>>

elders_cleaned_dataset_by_visit <- read_rds(paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))  %>% 
  mutate(cohort = factor(cohort, levels = c("2015", "2016", "2017", "2018", "2019"))) %>% 
  mutate(bp_dia = if_else(bp_dia < 40, 9999, bp_dia),
         sp02 = if_else(sp02 < 86, 9999, sp02),
         bp_sys = if_else(bp_dia == 9999 | is.na(bp_dia), 9999, bp_sys)) %>% 
  replace_with_na(replace = list(sp02 = 9999, bp_sys = 9999, bp_dia = 9999)) %>% 
  mutate(cpd = as.character(aveperday),
         cpd = as.numeric(cpd),
         ppd = cpd/20,
         age = as.character(age),
         age = as.numeric(age),
         agequit = as.character(agequit),
         agequit = as.numeric(agequit),
         agefirst = as.character(agefirst),
         agefirst = as.numeric(agefirst),
         agequit2 = ifelse(agequit < agefirst | is.na(agequit), age, agequit),
         smoke_years = agequit2 - agefirst,
         pack_years = smoke_years*ppd)
  left_join(treatments_blind, by = c("area", "home_id_num"))

@@

# Summary stats for character/factor data
```{r}
# summary data for manuscript tables
char_funct_total <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>%
  #filter(area == "NPT") %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  mutate(total_n = n()) %>% 
  group_by(area) %>% 
  mutate(area_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = n()/total_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(new_var)
demographics_summary
}

char_funct_treatment <- function(var, data = elders_cleaned_dataset_by_visit){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>% 
  #filter(area == "NPT") %>% 
  arrange(adult_id_char, winter_id, sampling_visit) %>% 
  mutate(new_var = if_else(is.na(new_var), lead(new_var), new_var)) %>% 
  distinct(adult_id_char, .keep_all = TRUE) %>% # change this to adult_id_char or home as necessary
  mutate(total_n = n()) %>% 
  group_by(treatment) %>% 
  mutate(treatment_n = n()) %>% 
  group_by(treatment, new_var) %>% 
  mutate(n = n(),
         percent = n()/treatment_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(treatment, new_var)
demographics_summary
}

# distinct by participant
char_funct_total("gender")
char_funct_treatment("gender")
char_funct_total("hispanic")
char_funct_treatment("hispanic")
char_funct_total("race")
char_funct_treatment("race")
char_funct_total("education")
char_funct_treatment("education")
char_funct_total("cohort")
char_funct_treatment("cohort")
char_funct_total("tobusenow")
char_funct_treatment("tobusenow")
char_funct_total("tobuseever")
char_funct_treatment("tobuseever")
char_funct_total("cigarspipes")
char_funct_treatment("cigarspipes")

@@

# distinct by participant
num_funct_total("age")
num_funct_treatment("age")
num_funct_total("pack_years")
num_funct_treatment("pack_years")

@@

# distinct by participant
num_funct_total("bp_sys")
num_funct_treatment("bp_sys")
num_funct_total("bp_dia")
num_funct_treatment("bp_dia")
num_funct_total("hr")
num_funct_treatment("hr")
num_funct_total("sp02")
num_funct_treatment("sp02")
num_funct_total("bmi")
num_funct_treatment("bmi")
num_funct_total("fvc")
num_funct_treatment("fvc")
num_funct_total("fev1")
num_funct_treatment("fev1")
num_funct_total("fev1_fvc_ratio")
num_funct_treatment("fev1_fvc_ratio")
num_funct_total("fvc_pp")
num_funct_treatment("fvc_pp")
num_funct_total("fev1_pp")
num_funct_treatment("fev1_pp")
num_funct_total("mp_mean_winter")
num_funct_treatment("mp_mean_winter")

@@

Health History Summary
```{r}
health_hx_data1 <- elders_cleaned_dataset_by_visit %>% 
  select(area, home, adult_id_char, winter_id, treatment, sampling_visit, age,
         rheum_arth:cigarspipes, residents_smoke, residents_smoke_inside) %>% 
  mutate_all(as.factor) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  distinct(area, adult_id_char, .keep_all = T) %>% 
  mutate(cpd = as.character(aveperday),
         cpd = as.numeric(cpd),
         ppd = cpd/20,
         age = as.character(age),
         age = as.numeric(age),
         agequit = as.character(agequit),
         agequit = as.numeric(agequit),
         agefirst = as.character(agefirst),
         agefirst = as.numeric(agefirst),
         agequit2 = ifelse(agequit < agefirst | is.na(agequit), age, agequit),
         smoke_years = agequit2 - agefirst,
         pack_years = smoke_years*ppd)

summary(health_hx_data)

health_hx_data_any <- health_hx_data1 %>% 
  select(area, home, adult_id_char, winter_id, sampling_visit,
         asthma, emphysema, chronicbronchitis, copd, hayfever, 
         pleuraleffusion, pleurisy, pneumonia, pneumothorax, coronaryarterydisease,
         angina, heartattack, chf, abnormalhr, pacemaker, icd, valvedisease,
         hypertension, stroke, diabetes, thyroiddisease, anemia, cancer) %>% 
  mutate(across(asthma:cancer, ~replace_na(.x, "No")))

summary(health_hx_data_any)

health_hx_data_cvd <- health_hx_data_any %>% 
  select(area, home, adult_id_char, winter_id, sampling_visit,
         coronaryarterydisease, angina, heartattack, chf, abnormalhr, pacemaker, 
         icd, valvedisease, hypertension, stroke, anemia)

summary(health_hx_data_cvd)

health_hx_data_ascvd <- health_hx_data_any %>% 
  select(area, home, adult_id_char, winter_id, sampling_visit,
         coronaryarterydisease, angina, heartattack, stroke)

summary(health_hx_data_ascvd)

health_hx_data_resp <- health_hx_data_any %>% 
  select(area, home, adult_id_char, winter_id, sampling_visit,
         asthma, emphysema, chronicbronchitis, copd, hayfever, 
         pleuraleffusion, pleurisy, pneumonia, pneumothorax)

summary(health_hx_data_resp)

health_hx_data <- elders_cleaned_dataset_by_visit %>% 
  select(area, home, adult_id_char, winter_id, treatment, sampling_visit,
         rheum_arth:cigarspipes, residents_smoke, residents_smoke_inside) %>% 
  mutate_all(as.factor) %>% 
  arrange(area, home, adult_id_char, winter_id, sampling_visit) %>% 
  distinct(area, adult_id_char, .keep_all = T) %>% 
  select(area, treatment, asthma, emphysema, chronicbronchitis, copd, hayfever, 
         pleuraleffusion, pleurisy, pneumonia, pneumothorax, coronaryarterydisease,
         angina, heartattack, chf, abnormalhr, pacemaker, icd, valvedisease,
         hypertension, stroke, diabetes, thyroiddisease, anemia, cancer) %>% 
  mutate(cvd_condition = as.factor(if_else(rowSums(health_hx_data_cvd == "Yes") > 0, "Yes", "No"))) %>% 
  mutate(ascvd_condition = as.factor(if_else(rowSums(health_hx_data_ascvd == "Yes") > 0, "Yes", "No"))) %>% 
  mutate(resp_condition = as.factor(if_else(rowSums(health_hx_data_resp == "Yes") > 0, "Yes", "No"))) %>% 
  mutate(any_condition = as.factor(if_else(rowSums(health_hx_data_any == "Yes") > 0, "Yes", "No")))

summary(health_hx_data)
```

```{r}
# summary data for manuscript tables
char_funct_total <- function(var, data = health_hx_data){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>%
  #filter(area == "NPT") %>% 
  mutate(total_n = n()) %>% 
  group_by(area) %>% 
  mutate(area_n = n()) %>% 
  group_by(new_var) %>% 
  mutate(n = n(),
         percent = n()/total_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(new_var)
demographics_summary
}

char_funct_treatment <- function(var, data = health_hx_data){
demographics_summary <- data %>% 
  rename(new_var = var) %>% 
  #filter(treatment_assigned == "Placebo") %>% 
  #filter(area == "NPT") %>% 
  mutate(total_n = n()) %>% 
  group_by(treatment) %>% 
  mutate(treatment_n = n()) %>% 
  group_by(treatment, new_var) %>% 
  mutate(n = n(),
         percent = n()/treatment_n) %>% 
  distinct(new_var, .keep_all = TRUE) %>% 
  dplyr::select(new_var, n, percent) %>% 
  arrange(treatment, new_var)
demographics_summary
}

# distinct by participant
char_funct_total("asthma")
char_funct_treatment("asthma")
char_funct_total("chronicbronchitis")
char_funct_treatment("chronicbronchitis")
char_funct_total("cancer")
char_funct_treatment("cancer")
char_funct_total("coronaryarterydisease")
char_funct_treatment("coronaryarterydisease")
char_funct_total("chf")
char_funct_treatment("chf")
char_funct_total("copd")
char_funct_treatment("copd")
char_funct_total("diabetes")
char_funct_treatment("diabetes")
char_funct_total("emphysema")
char_funct_treatment("emphysema")
char_funct_total("hayfever")
char_funct_treatment("hayfever")
char_funct_total("heartattack")
char_funct_treatment("heartattack")
char_funct_total("hypertension")
char_funct_treatment("hypertension")
char_funct_total("stroke")
char_funct_treatment("stroke")
char_funct_total("cvd_condition")
char_funct_treatment("cvd_condition")
char_funct_total("ascvd_condition")
char_funct_treatment("ascvd_condition")
char_funct_total("resp_condition")
char_funct_treatment("resp_condition")
char_funct_total("any_condition")
char_funct_treatment("any_condition")
```
